Hierarchical Dispatching and Overriding

### 2017-09-30 NewVersion/sections/Overview.tex
Section 2.1

\section{A Running Example: DrawableDeck}~\label{sec:overview}
This section illustrates the features of our model for
resolving unintentional method conflicts. As mentioned before, such a
case arises when two inherited methods happen to have the same
signature, but with different semantics and functionality. This
situation is troublesome to programmers that use multiple
inheritance. Below we illustrate with a running example called
\lstinline|DrawableDeck|, which models a drawable deck of cards. 
Note that we use a Java-like syntax
throughout the paper, and all types are defined with the keyword
``\lstinline|interface|''. The concept is closely related to Java 8
interfaces with default methods~\cite{bono14} and traits. For simplicity, an interface in our model has
the following characteristics:
\begin{itemize}
  \item It allows multiple inheritance.
  \item Every method represents a behaviour and requires a body for its implementation (like Java 8 default methods). There are no abstract methods.
  \item It can be directly instantiated by the \lstinline|new| keyword.
\end{itemize}
Furthermore for simplicity, our examples and formalization do not deal with state at
this stage. %%, but that should not block our discussions below.

@@

\vspace{3pt}\begin{lstlisting}
interface Deck {
  void draw() { // draws a card from the Deck
    Stack<Card> cards = this.getStack();
    if (!cards.isEmpty()) { Card card = cards.pop(); ... }
  }
}
interface Drawable {
  void draw() { // draws something on the screen
    JFrame frame = new JFrame("Canvas");
    frame.setVisible(true);
    ...
  }
}
\end{lstlisting}\vspace{3pt}
Note that both methods have \lstinline|void| return type (we will not formalize
\lstinline|void| in our calculus afterwards; here it is only for illustration). 
In \lstinline|Deck|, \lstinline|draw| tries to get the cards as a stack, pops
out the top card, and so on. While in \lstinline|Drawable|, \lstinline|draw|
creates a blank canvas using \lstinline|JFrame|. Now, a programmer is designing a
card game with GUI. He may want to draw a deck on the screen, so he defines a drawable
deck using multiple inheritance:

@@

Now one may quickly come up with a workaround, which is creating a new \lstinline|draw| method in \lstinline|DrawableDeck| to
explicitly override the old ones. However, the two conflicting methods have totally different functionalities, merging
them into one does not make any sense. This non-solution will hide the old methods and break independent extensibility. The
feature we expect is called ``triangle inheritance'', that is to say, when the conflicting methods serve for different behaviours,
we can allow them to coexist in the extending interface in a type-safe way.\\

\noindent\textbf{Potential Fixes} There are several workarounds that come to our mind:
\begin{itemize}
  \item \textbf{I. Delegation.} As an alternative to multiple inheritance,
  delegation can be used by introducing two fields with
  \lstinline|Drawable| type and \lstinline|Deck| type,
  respectively. This avoids method conflicts. Nevertheless, it is known
  that using delegation makes it hard to correctly maintain
  self-references in an extensible system and also
  introduces a lot of boilerplate code.
  \item \textbf{II. Refactor Drawable and/or Deck to rename the methods.} If
  the source code for \lstinline|Drawable| or \lstinline|Deck| is available
  then it may be possible to rename one of the \lstinline|draw|
  methods. However this approach is non-modular, as it requires 
  modifying existing code, and may not be possible if code is unavailable.
  \item \textbf{III. Method exclusion/renaming in traits.} Some trait models
  support method exclusion/renaming. Those features
   can eliminate conflicts, although most
  programming languages do not support them. In a traditional OO system,
  they can break the subtyping relationship. Moreover, in
  contrast with exclusion, renaming can indeed preserve both conflicting
  behaviours, however, it is cumbersome in practice, as introducing new
  names can affect other code blocks.
  \item \textbf{IV. Static dispatch.} Some other languages like C++ have
  direct support for the triangle inheritance. While conflicting methods are
  both preserved, the ambiguity of method call is resolved by static dispatch.
  Qualified names are used to specify the branch to give the right implementation.
\end{itemize}

\noindent\textbf{Our solution:} below is our code that shows the triangle
inheritance.
We can define \lstinline|DrawableDeck| by inheriting \lstinline|Drawable| and \lstinline|Deck| together while
type safety is ensured. Again it is slightly different from our prototype as we do not model \lstinline|void| or statements;
the code is just for illustration.
\vspace{3pt}\begin{lstlisting}
interface DrawableDeck extends Drawable, Deck {}
interface Foo {
  void func(Deck obj) { obj.draw(); }
}
// main program
DrawableDeck d = new DrawableDeck();
new DrawableDeck().Drawable::draw();      // calling draw in Drawable
new Foo().func(new DrawableDeck());       // calling draw in Deck
\end{lstlisting}\vspace{3pt}
Note that both cases in the above code are valid in C++ for disambiguation. As shown by \lstinline|func|, C++ looks at
the static type of the receiver for static dispatching. Normally we can define such a method \lstinline|func| for ``implicit''
casts in \name{}. C++ further allows explicit casts.

In the following text, we will see that C++ is still unsatisfactory, as it introduces another problem on code reuse, which motivates
our work.

>>>>

\section{A Running Example: DrawableDeck}~\label{sec:overview}
This section illustrates the features of our model for
resolving unintentional method conflicts. As mentioned before, such a
case arises when two inherited methods happen to have the same
signature, but with different semantics and functionality. This
situation is troublesome for programmers that use multiple
inheritance. Below we illustrate with a running example called
\lstinline|DrawableDeck|, which models a drawable deck of cards. 
Note that we use a Java-like syntax
throughout the paper, and all types are defined with the keyword
``\lstinline|interface|''. The concept is closely related to Java 8
interfaces with default methods~\cite{bono14} and traits. For simplicity, an interface in our model has
the following characteristics:
\begin{itemize}
  \item It allows multiple inheritance.
  \item Every method represents a behaviour and requires a body for
    its implementation (like Java 8 default methods). 
  \item It can be directly instantiated by the \lstinline|new| keyword.
\end{itemize}
Furthermore, similarly to traits or Java 8 with default methods,
interfaces cannot have state.\bruno{I rephrased this sentence}

%%for simplicity, our examples and formalization do not deal with state at
%%this stage. %%, but that should not block our discussions below.

@@

\vspace{3pt}\begin{lstlisting}
interface Deck {
  void draw() { // draws a card from the Deck
    Stack<Card> cards = this.getStack();
    if (!cards.isEmpty()) { Card card = cards.pop(); ... }
  }
}
interface Drawable {
  void draw() { // draws something on the screen
    JFrame frame = new JFrame("Canvas");
    frame.setVisible(true);
    ...
  }
}
\end{lstlisting}\vspace{3pt}
\bruno{The code calls getStack(), but that method is not defined! 
If you want to present a compelling example, please make sure that 
the code makes sense! I suppose that getStack() would be an abstract method?
}
Note that both methods have \lstinline|void| return type (we will not formalize
\lstinline|void| in our calculus afterwards; here it is only for illustration). 
In \lstinline|Deck|, \lstinline|draw| tries to get the cards as a stack, pops
out the top card, and so on. While in \lstinline|Drawable|, \lstinline|draw|
creates a blank canvas using \lstinline|JFrame|. Now, a programmer is designing a
card game with GUI. He may want to draw a deck on the screen, so he defines a drawable
deck using multiple inheritance:

@@

Now one may quickly come up with a workaround, which is creating a new \lstinline|draw| method in \lstinline|DrawableDeck| to
explicitly override the old ones. However, the two conflicting methods have totally different functionalities, merging
them into one does not make any sense. This non-solution will hide the
old methods and break independent extensibility.

\noindent\textbf{Potential Fixes} There are several other workarounds
that come to our mind. We briefly discuss those potential fixes and
workarounds next:
\begin{itemize}
  \item \textbf{I. Delegation.} As an alternative to multiple inheritance,
  delegation can be used by introducing two fields with
  \lstinline|Drawable| type and \lstinline|Deck| type,
  respectively. This avoids method conflicts. Nevertheless, it is known
  that using delegation makes it hard to correctly maintain
  self-references in an extensible system and also
  introduces a lot of boilerplate code.
  \item \textbf{II. Refactor Drawable and/or Deck to rename the methods.} If
  the source code for \lstinline|Drawable| or \lstinline|Deck| is available
  then it may be possible to rename one of the \lstinline|draw|
  methods. However this approach is non-modular, as it requires 
  modifying existing code, and may not be possible if code is unavailable.
  \item \textbf{III. Method exclusion/renaming in traits.} Some trait models
  support method exclusion/renaming. Those features
   can eliminate conflicts, although most
  programming languages do not support them. In a traditional OO system,
  they can break the subtyping relationship. Moreover, in
  contrast with exclusion, renaming can indeed preserve both conflicting
  behaviours, however, it is cumbersome in practice, as introducing new
  names can affect other code blocks.
  \item \textbf{IV. Static dispatch.} Some other languages like C++ have
  direct support for the triangle inheritance. While conflicting methods are
  both preserved, the ambiguity of method call is resolved by static dispatch.
  Qualified names are used to specify the branch to give the right
  implementation.\bruno{don't forget to readjust text with what we
    have learned about C++.}
\end{itemize}

\noindent\textbf{Our solution:} 
The feature we expect is called \emph{triangle inheritance}, that is
to say, when the conflicting methods serve for different behaviours,
we can allow them to coexist in the extending interface in a type-safe way.
Below is our code that shows the triangle
inheritance.
We can define \lstinline|DrawableDeck| by inheriting \lstinline|Drawable| and \lstinline|Deck| together while
type safety is ensured. 
%%Again it is slightly different from our prototype as we do not model \lstinline|void| or statements;
%%the code is just for illustration.
\vspace{3pt}\begin{lstlisting}
interface DrawableDeck extends Drawable, Deck {}
interface Foo {
  void func(Deck obj) { obj.draw(); }
}
// main program
DrawableDeck d = new DrawableDeck();
new DrawableDeck().Drawable::draw();      // calling draw in Drawable
new Foo().func(new DrawableDeck());       // calling draw in Deck
\end{lstlisting}\vspace{3pt}
\bruno{Please ensure that the syntax we use in the examples, matches 
up with the syntax in the formalization. In particular make sure that
the syntax for the various types of method calls are the same.}
\bruno{You don't actually explain your code! Please explain what the
  code does, and what are the different types of calls
  involved. Introduce some terminology here, namely the names for the 
different types of calls.}

\noindent\textbf{A comparison with C++:} 
Note that both cases in the above code are valid in C++ for disambiguation. As shown by \lstinline|func|, C++ looks at
the static type of the receiver for static dispatching. Normally we can define such a method \lstinline|func| for ``implicit''
casts in \name{}. C++ further allows explicit casts.
In the following text, we will see that C++ is still unsatisfactory, as it introduces another problem on code reuse, which motivates
our work.\bruno{Please revise.}
